### 2019-01-02 example programs/goEnrichment.GSTA4KO.2017-09-01.Rmd
add example programs

---
title: "GSTA4 KO analysis for Colin"
author: "Laura Saba"
date: "9/1/2017"
output: html_document
---

```{r}
rm(list=ls())
options(stringsAsFactors = FALSE)
library("GO.db")
library(dplyr)
wd <- "~/Documents/Shearn/GSTA4/"

orig <- read.table(file=paste(wd,"data/120d EtOH GSTA4 study Table for Laura Saba.txt",sep=""),sep="\t",quote="",header=TRUE,fill=TRUE)
orig <- orig[orig$Accession!="totals" & orig$Accession!=" " & orig$Accession!="",]

update <- read.table(file=paste(wd,"data/updateProteinID.2017-07-14.txt",sep=""),sep="\t",header=TRUE)
proteins <- merge(orig,update,by.x="Accession",by.y="orig",all=TRUE)
proteins$Accession[!is.na(proteins$new)] = proteins$new[!is.na(proteins$new)]

# GO data base
go_db <- read.table(file=paste(wd,"source/",sep=""),sep="\t",header=FALSE,comment.char="!",quote="")

library(biomaRt)
ensembl=useMart("ensembl")
ensembl = useDataset("mmusculus_gene_ensembl",mart=ensembl)

#convert protein ID to Uniprot symbols
protein_to_uniprot <- read.table(file=paste(wd,"data/protein_to_uniprot.2017-07-14.txt",sep=""),sep="\t",quote="",header=TRUE,fill=TRUE)

proteins <- merge(proteins,protein_to_uniprot,by.x="Accession",by.y="Entry.name",all.x=TRUE)

##problematic proteins
proteins <- proteins[!(proteins$Accession %in% c("PZP_MOUSE","SLF2_MOUSE")),]

##Convert to MGI Symbol
toMGI <- getBM(attributes=c('uniprot_gn', 'mgi_id',"mgi_symbol"), 
      filters = 'mgi_id', 
      values = go_db$V2, 
      mart = ensembl)
```

GO Enrichment

```{r}
uniprot_to_go <- merge(toMGI,go_db,by.x="mgi_id",by.y="V2")

Chow_GSTA4 <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$Chow.GSTA4...==1]])
Chow_SV <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$Chow.SV==1]])
EtOH_GSTA4 <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$EtOH.GSTA4...==1]])
EtOH_SV <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$EtOH.SV==1]])
Binge_GSTA4 <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$Binge.GSTA4...==1]])
Binge_SV <- unique(uniprot_to_go$uniprot_gn[uniprot_to_go$uniprot_gn %in% proteins$Entry[proteins$Binge.SV==1]])

pathEnrich <- function(x){
  all_GO <- unique(uniprot_to_go$uniprot_gn)
  sig_GO <- unique(x)

  all_GO_cnt <- uniprot_to_go[uniprot_to_go$uniprot_gn %in% all_GO,]
  sig_GO_cnt <- uniprot_to_go[uniprot_to_go$uniprot_gn %in% sig_GO,]

  all_GO_cnt <- all_GO_cnt %>%
    group_by(V5) %>%
    summarize(GO_cnt = length(uniprot_gn))

  sig_GO_cnt <- sig_GO_cnt %>%
    group_by(V5) %>%
    summarize(GO_sig = length(uniprot_gn))

  enrich_table <- merge(all_GO_cnt,sig_GO_cnt,by="V5",all=TRUE)
  enrich_table$GO_sig[is.na(enrich_table$GO_sig)] = 0

  enrich_table$numTested <- length(all_GO)
  enrich_table$numSig <- length(sig_GO)
  enrich_table <- enrich_table %>%
    mutate(expected = (numSig/numTested)*GO_cnt)

  #enrich_table <- merge(path_names_mmu,enrich_table,by.x="V1",by.y="pathway")

  enrich_table$enrich_p <- NA
  for(i in 1:nrow(enrich_table)){
    a = enrich_table[i,"GO_sig"]
    b = enrich_table[i,"GO_cnt"] - enrich_table[i,"GO_sig"]
    c = enrich_table[i,"numSig"] - enrich_table[i,"GO_sig"]
    d = enrich_table[i,"numTested"] - enrich_table[i,"numSig"] + enrich_table[i,"GO_sig"]
    enrich_table$enrich_p[i] = fisher.test(matrix(c(a,b,c,d),nr=2),alternative="greater")$p.value
  }

  enrich_table <- enrich_table[order(enrich_table$enrich_p),]
  enrich_table$fdr <- p.adjust(enrich_table$enrich_p,method="BH")
  return(enrich_table)
}

Chow_GSTA4_enrich <- pathEnrich(Chow_GSTA4)
Chow_SV_enrich <- pathEnrich(Chow_SV)
EtOH_GSTA4_enrich <- pathEnrich(EtOH_GSTA4)
EtOH_SV_enrich <- pathEnrich(EtOH_SV)
Binge_GSTA4_enrich <- pathEnrich(Binge_GSTA4)
Binge_SV_enrich <- pathEnrich(Binge_SV)

colnames(Chow_GSTA4_enrich) <- c("V1","GO_cnt","GO_sig_Chow_GSTA4","numTested","numSig_Chow_GSTA4","expected_Chow_GSTA4","enrich_p_Chow_GSTA4","fdr_Chow_GSTA4")
colnames(Chow_SV_enrich) <- c("V1","GO_cnt","GO_sig_Chow_SV","numTested","numSig_Chow_SV","expected_Chow_SV","enrich_p_Chow_SV","fdr_Chow_SV")

colnames(EtOH_GSTA4_enrich) <- c("V1","GO_cnt","GO_sig_EtOH_GSTA4","numTested","numSig_EtOH_GSTA4","expected_EtOH_GSTA4","enrich_p_EtOH_GSTA4","fdr_EtOH_GSTA4")
colnames(EtOH_SV_enrich) <- c("V1","GO_cnt","GO_sig_EtOH_SV","numTested","numSig_EtOH_SV","expected_EtOH_SV","enrich_p_EtOH_SV","fdr_EtOH_SV")

colnames(Binge_GSTA4_enrich) <- c("V1","GO_cnt","GO_sig_Binge_GSTA4","numTested","numSig_Binge_GSTA4","expected_Binge_GSTA4","enrich_p_Binge_GSTA4","fdr_Binge_GSTA4")
colnames(Binge_SV_enrich) <- c("V1","GO_cnt","GO_sig_Binge_SV","numTested","numSig_Binge_SV","expected_Binge_SV","enrich_p_Binge_SV","fdr_Binge_SV")

combined_enrich <- merge(Chow_GSTA4_enrich,Chow_SV_enrich,by=c("V1","GO_cnt","numTested"),all=TRUE)
combined_enrich <- merge(combined_enrich,EtOH_GSTA4_enrich,by=c("V1","GO_cnt","numTested"),all=TRUE)
combined_enrich <- merge(combined_enrich,EtOH_SV_enrich,by=c("V1","GO_cnt","numTested"),all=TRUE)
combined_enrich <- merge(combined_enrich,Binge_GSTA4_enrich,by=c("V1","GO_cnt","numTested"),all=TRUE)
combined_enrich <- merge(combined_enrich,Binge_SV_enrich,by=c("V1","GO_cnt","numTested"),all=TRUE)

threshold = 0.01
sum(combined_enrich$fdr_Chow_GSTA4<threshold,na.rm=TRUE)
sum(combined_enrich$fdr_Chow_SV<threshold,na.rm=TRUE)
sum(combined_enrich$fdr_EtOH_GSTA4<threshold,na.rm=TRUE)
sum(combined_enrich$fdr_EtOH_SV<threshold,na.rm=TRUE)
sum(combined_enrich$fdr_Binge_GSTA4<threshold,na.rm=TRUE)
sum(combined_enrich$fdr_Binge_SV<threshold,na.rm=TRUE)

combined_enrich$num_grps_sig <- rowSums(combined_enrich[,grep("fdr",colnames(combined_enrich))]<threshold)

of_interest <- combined_enrich[combined_enrich$num_grps_sig>0 & combined_enrich$num_grps_sig<6,]

groups <- paste(rep(c("Chow","EtOH","Binge"),each=2),rep(c("GSTA4","SV"),3),sep="_")
of_interest[,c("V1","V2",paste("fdr_",groups,sep=""))]

combined_enrich$sig_pattern <- apply(combined_enrich[,grep("fdr",colnames(combined_enrich))],1,function(a) paste(as.numeric(a<threshold),collapse=""))
table(rowSums(combined_enrich[,grep("enrich_p",colnames(combined_enrich))]<threshold))
```

##Differential Enrichment

```{r}
diffEnrich <- function(a,b,c,d,OR){
  y = fisher.test(matrix(c(a,b,c,d),nr=2))
  if(OR) z = y$estimate
  if(!OR) z = y$p.value
  return(z)
}

#diff enrich EtOH effect - WT Chow vs WT EtOH

combined_enrich$EtOH_Effect_p = apply(combined_enrich[,c("GO_sig_Chow_SV","GO_sig_EtOH_SV","numSig_Chow_SV","numSig_EtOH_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$EtOH_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_Chow_SV","GO_sig_EtOH_SV","numSig_Chow_SV","numSig_EtOH_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich EtOH effect in GSTA4 - GSTA4 Chow vs GSTA4 EtOH

combined_enrich$EtOHinGSTA4_Effect_p = apply(combined_enrich[,c("GO_sig_Chow_GSTA4","GO_sig_EtOH_GSTA4","numSig_Chow_GSTA4","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$EtOHinGSTA4_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_Chow_GSTA4","GO_sig_EtOH_GSTA4","numSig_Chow_GSTA4","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Binge effect - WT Binge vs WT EtOH

combined_enrich$Binge_Effect_p = apply(combined_enrich[,c("GO_sig_EtOH_SV","GO_sig_Binge_SV","numSig_EtOH_SV","numSig_Binge_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$Binge_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_EtOH_SV","GO_sig_Binge_SV","numSig_EtOH_SV","numSig_Binge_SV")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Binge effect in GSTA4 - GSTA4 Binge vs GSTA4 EtOH

combined_enrich$BingeInGSTA4_Effect_p = apply(combined_enrich[,c("GO_sig_EtOH_GSTA4","GO_sig_Binge_GSTA4","numSig_EtOH_GSTA4","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$BingeInGSTA4_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_EtOH_GSTA4","GO_sig_Binge_GSTA4","numSig_EtOH_GSTA4","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Genotype Effect - WT Chow vs GSTA4 Chow

combined_enrich$Genotype_Effect_p = apply(combined_enrich[,c("GO_sig_Chow_SV","GO_sig_Chow_GSTA4","numSig_Chow_SV","numSig_Chow_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$Genotype_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_Chow_SV","GO_sig_Chow_GSTA4","numSig_Chow_SV","numSig_Chow_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

#diff enrich Genotype Effect With Ethanol - WT EtOH vs GSTA4 EtOH

combined_enrich$GenoInEtOH_Effect_p = apply(combined_enrich[,c("GO_sig_EtOH_SV","GO_sig_EtOH_GSTA4","numSig_EtOH_SV","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$GenoInEtOH_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_EtOH_SV","GO_sig_EtOH_GSTA4","numSig_EtOH_SV","numSig_EtOH_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

combined_enrich[combined_enrich$GenoInEtOH_Effect_p<0.05,]

#diff enrich Genotype Effect With Binge - WT Binge vs GSTA4 Binge

combined_enrich$GenoInBinge_Effect_p = apply(combined_enrich[,c("GO_sig_Binge_SV","GO_sig_Binge_GSTA4","numSig_Binge_SV","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=FALSE))

combined_enrich$GenoInBinge_Effect_OddsRatio = apply(combined_enrich[,c("GO_sig_Binge_SV","GO_sig_Binge_GSTA4","numSig_Binge_SV","numSig_Binge_GSTA4")],1,function(a) diffEnrich(a[1],a[2],a[3],a[4],OR=TRUE))

table(rowSums(combined_enrich[,grep("_Effect_p",colnames(combined_enrich),fixed=TRUE)]<0.05))

### add FDR  ##
combined_enrich$GenoInBinge_Effect_fdr = p.adjust(combined_enrich$GenoInBinge_Effect_p,method="BH")
combined_enrich$GenoInEtOH_Effect_fdr = p.adjust(combined_enrich$GenoInEtOH_Effect_p,method="BH")
combined_enrich$Genotype_Effect_fdr = p.adjust(combined_enrich$Genotype_Effect_p,method="BH")
combined_enrich$BingeInGSTA4_Effect_fdr = p.adjust(combined_enrich$BingeInGSTA4_Effect_p,method="BH")
combined_enrich$Binge_Effect_fdr = p.adjust(combined_enrich$Binge_Effect_p,method="BH")
combined_enrich$EtOHinGSTA4_Effect_fdr = p.adjust(combined_enrich$EtOHinGSTA4_Effect_p,method="BH")
combined_enrich$EtOH_Effect_fdr = p.adjust(combined_enrich$EtOH_Effect_p,method="BH")

tmp_results <- combined_enrich[rowSums(combined_enrich[,grep("_Effect_p",colnames(combined_enrich),fixed=TRUE)]<0.05)>0,]

xx<-as.list(GOTERM)
y <- xx["GO:0002818"]]

anno <- do.call('rbind',lapply(xx[tmp_results$V1],function(y) data.frame(GO_ID = GOID(y),
           Term = Term(y),
           Onto = Ontology(y),
           Defin = Definition(y))))

tmp_results <- merge(anno,tmp_results,by.x="GO_ID",by.y="V1")
write.table(tmp_results,file="~/Desktop/tmp_results_2017-09-07.txt",sep="\t",row.names = FALSE,col.names = TRUE)
```